Wigglyhedra

### 2024-04-08 draft/wigglyhedra.tex
Added some categorical context. Needs to be cleaned up.

% operators
\DeclareMathOperator{\conv}{conv} % convex hull
\DeclareMathOperator{\vect}{vect} % linear span
\DeclareMathOperator{\cone}{cone} % cone hull
\DeclareMathOperator{\inv}{inv} % inversions
\DeclareMathOperator{\ninv}{ninv} % inversions
\DeclareMathOperator{\Ima}{Im} % image
\DeclareMathOperator{\Vol}{Vol} % (mixed) volume

@@

Analogies to cluster algebras and gentle algebras.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

>>>>

% operators
\DeclareMathOperator{\conv}{conv} % convex hull
\DeclareMathOperator{\vect}{vect} % linear span
\DeclareMathOperator{\cone}{cone} % cone hull
\DeclareMathOperator{\inv}{inv} % inversions
\DeclareMathOperator{\ninv}{ninv} % inversions
\DeclareMathOperator{\Ima}{Im} % image
\DeclareMathOperator{\Vol}{Vol} % (mixed) volume
\DeclareMathOperator{\Hom}{Hom} % hom-spaces

@@

Analogies to cluster algebras and gentle algebras.

\subsection{Curves in a disk and a triangulated category associated to a quiver}
There is an algebraic/categorical model of curves in a disk with marked points, due to Khovanov and Seidel~\cite{kho.sei:02}, which we describe in this section.
This model identifies curves in a disk with objects in a certain triangulated category, and the correspondence has several useful features.
For instance, the number of intersections between two curves is the number of morphisms in the category between the corresponding objects.
Furthermore, the Artin braid group is the fundamental group of the unordered configuration space of points in a disk, and thus acts on the set of curves in the disk.
The associated triangulated category also carries an action by the Artin braid group, and the Khovanov--Seidel map from curves to objects is equivariant with respect to this group action.

We now give a concise account of this correspondence.
We omit the details that do not play a starring role in the remainder of this paper, and direct the reader to the original references.
Recall that the \emph{type \(A_m\)} Dynkin graph is an undirected graph with \(m\) vertices and \(m-1\) edges; each edge connecting successive pairs of points as shown in~\cref{fig:a5-dynkin}.
\begin{figure}[h]
  \centering
  \begin{tikzcd}
    \bullet \arrow[-]{r}& \bullet \arrow[-]{r}& \bullet \arrow[-]{r}& \bullet \arrow[-]{r}& \bullet
  \end{tikzcd}
  \caption{The type \(A_5\) Dynkin graph.}
  \label{fig:a5-dynkin}
\end{figure}
Given an undirected graph as above, we can construct its \emph{doubled quiver}, which is a directed graph in which every undirected edge is replaced by a pair of oppositely oriented arrows, as shown in~\cref{fig:a5-doubled}.
\begin{figure}[h]
  \centering
  \begin{tikzcd}
    \bullet \arrow[bend left]{r}& \bullet \arrow[bend left]{r} \arrow[bend left]{l}& \bullet \arrow[bend left]{r}\arrow[bend left]{l}& \bullet \arrow[bend left]{r}\arrow[bend left]{l}& \bullet \arrow[bend left]{l}
  \end{tikzcd}
  \caption{The doubled quiver of the type \(A_5\) Dynkin graph.}
  \label{fig:a5-doubled}
\end{figure}
The \emph{zig-zag algebra} of the doubled quiver of an undirected graph is a finite-dimensional quotient of the path algebra of the doubled quiver of that graph (see~\cite[Section 3]{hue.kho:01} for the precise definition).

Then there is a triangulated category \(\mathcal{C}_m\) associated to the Dynkin graph of type \(A_m\), constructed as a differential graded version of the homotopy category of projective modules over the zig-zag algebra of the doubled quiver of type \(A_m\).
The precise definition of this category appears in, e.g.,~\cite{sei.tho:01,tho:06} for type \(A_m\) and in~\cite{bap.deo.lic:20} for general undirected graphs.

The category \(\mathcal{C}_m\) is linear over a fixed field \(k\), triangulated, and 2-Calabi--Yau.
That is, for every pair of objects \(X\) and \(Y\) in \(\mathcal{C}_m\), we have
\[\Hom_{\mathcal{C}_m}(X,Y[d]) \cong \Hom_{\mathcal{C}_m}(Y, X[2-d]),\]
where \([1]\) denotes the triangulated shift.
\asilata{Update these definitions to the bigraded version, not the 2-CY version.}
Furthermore, \(\mathcal{C}_m\) is strongly generated by \(m\) objects denoted by \(P_1, \ldots, P_m\), which correspond to indecomposable projective modules over the zig-zag algebra.
These objects satisfy:
\begin{itemize}
\item \(\Hom(P_i,P_i[d]) = 0\) if \(d \neq 0, 2\);
\item \(\Hom(P_i,P_i[d]) = k\) if \(d = 0, 2\);
\item \(\Hom(P_i, P_{i\pm1}[d]) = 0\) if \(d \neq 1\);
\item \(\Hom(P_i, P_{i\pm1}[1]) = k\), and
\item \(\Hom(P_i, P_j[d]) = 0 \) otherwise.
\end{itemize}

Now consider \(n+2\) marked points on a disk, arranged on a straight horizontal line, and labelled \(0,1, \ldots, n+1\).
We will consider isotopy classes of curves whose endpoints lie on distinct marked points, and which do not cross themselves or any marked point in their interior.
For convenience, we also refer to such an isotopy class as a \emph{curve}.
The Khovanov--Seidel correspondence (explained in Section 4 of~\cite{kho.sei:02}) associates objects of \(\mathcal{C}_{n+1}\) to curves on a configuration of \(n+2\) marked points arranged in a straight line.

In particular, the line segment joining point \(i\) to point \(i+1\) is sent to the object \(P_{i+1}\) in the category \(\mathcal{C}_{n+1}\).
It is easy to check that the (minimal) intersection numbers between any two such segments counts the number of morphisms between the associated objects.
In particular, the self-intersection number of a single segment is computed to be \(2\), as the second copy of the segment can be slightly deformed to be disjoint from the first copy except at the two endpoints.
In fact, it is true for any pair of curves that its minimal intersection number is the number of morphisms between the associated objects (\cite[Lemma 4.11]{kho.sei:02}).
\asilata{Now figure out what to say about \(t\)-structures, namely that wiggly arcs are exactly those whose associated complexes lie in the heart of the standard \(t\)-structure.}
